Greenhouse gas emissions model

### 2021-08-23 210617_GHG_model.Rmd
added herd functionality

```{r make_variables}
###CW Note ####
### We will need to clean up the warnings later
### Reasonable ranges still missing for many variables
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}

@@

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #####in these code sections I am missing the overview. 
  ##### it is not immediately clear what this is meant to do... 
  ##### from a specific publication?

# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
#####EL CW Note ####
## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.

@@

# We're not sure what the animal types stand for, but this is important for calculating emissions.
# We need to know, which types are dairy cattle, which ones are lactating cows, which ones are bulls, and which ones are 'others'. We have 12 animal types.

animal_types_lact_cow<-c(4,5,6) # lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant. This resulted in a weighted mean of the various maintenance coefficients. I'm not sure where we could get such data.

# animal type 1 - 'intact' grown-up males >=36 months
# animal type 2 - castrated grown-up males >=36 months
# animal type 3 - males 12-36 months
# animal type 4 - cows
# animal type 5 - cows
# animal type 6 - cows
# animal type 7 - females <= 12 months
# animal type 8 - males <= 12 months
# animal type 9 - females 12-36 months
# animal type 10 - female calves (<12 months)
# animal type 11 - male calves (<12 months)
# animal type 12 - first-time pregnant cows

# define C and Cfi based on sex (Page 10.17, below Equation 10.6)
if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)

@@

@@

# I don't quite understand how Ca was assigned in the original data table. There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36. I can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.

@@

#variables:
#milk_yield (kg milk per day)
#milk_fat (percent)

@@

# Ratio of net energy available for growth in a diet to digestable energy consumed (Equation 10.15)
REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)

@@

# Compute the CO2-equivalent of methane emissions for enteric methane
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CH4_CO2eq<-enteric_CH4*25

@@

# Estimate CO2eq for manure methane
mm_CH4_CO2eq<-mm_ch4*25

@@

# Estimate CO2eq for manure nitrous
mm_N2O_CO2eq<-(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            milk_yield=milk_yield,
            enteric_CH4_CO2eq=enteric_CH4_CO2eq,
            mm_CH4_CO2eq=mm_CH4_CO2eq,
            mm_N2O_CO2eq=mm_N2O_CO2eq,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

@@

GHG_simulation <- mcSimulation(
  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
  model_function = ghg_emissions,
  numberOfModelRuns = 1e2, #100 for now
  functionSyntax = "plainNames"
)

@@

plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Methane Emissions from Enteric Fermentation")

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4_CO2eq)
```

@@

Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4`.

```{r name_in_out_vars}
in_var  <- GHG_simulation$x$milk_yield
out_var <- GHG_simulation$y$enteric_CH4
```

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = 1, max_in_var = 4)
```

>>>>

```{r make_variables}
### Reasonable ranges still missing for many variables
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}

@@

# Enteric fermentation emissions ($CH_4$)

# The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use). The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).

@@

# We're not sure what the animal types stand for, but we inferred some information from the numbers in the sample dataset. We further looked for guidance in the "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" (https://www.ipcc-nggip.iges.or.jp/public/gp/bgp/4_1_CH4_Enteric_Fermentation.pdf)

# animal type 1 - 'intact' grown-up males >=36 months
# animal type 2 - castrated grown-up males >=36 months
# animal type 3 - males 12-36 months
# animal type 4 - lactating cows
# animal type 5 - lactating cows
# animal type 6 - lactating cows
# animal type 7 - weaned young females <= 12 months
# animal type 8 - weaned young males <= 12 months
# animal type 9 - females 12-36 months
# animal type 10 - pre-weaned female calves (<12 months)
# animal type 11 - pre-weaned male calves (<12 months)
# animal type 12 - first-time pregnant cows

# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant. This resulted in a weighted mean of the various maintenance coefficients. We're not sure where we could get such data.

  mm_N2O_CO2eq<-c()
  mm_CH4_CO2eq<-c()
  enteric_CH4_CO2eq<-c()
  total_milk<-c()

for (animal_type in 1:12)
{
# define C and Cfi based on sex (Page 10.17, below Equation 10.6)
if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)

@@

N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))

@@

# We don't quite understand how Ca was assigned in the original data table. There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36. We can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table. For now, we're restricting these Ca values to only those that are given as 'pure' categories in the IPCC guidelines (Table 10.5).

@@

#milk_yield (kg milk per day)
#milk_fat (percent)

@@

# Ratio of net energy available for growth in a diet to digestible energy consumed (Equation 10.15)
REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)

@@

# Compute the CO2-equivalent of methane emissions for enteric methane
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25

@@

# Estimate CO2eq for manure methane
mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25

@@

# Estimate CO2eq for manure nitrous
mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)

total_milk[animal_type]<-N_animal_type*milk_yield*365

}

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            milk_yield=sum(total_milk),
            enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
            mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
            mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
            feed_CO2=feed_CO2,
            on_farm=sum(on_farm)))
}

@@

GHG_simulation <- mcSimulation(
  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
  model_function = ghg_emissions,
  numberOfModelRuns = 1e3, #1000 for now
  functionSyntax = "plainNames"
)

@@

plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4_CO2eq",
                   method = "boxplot_density",
                   old_names = "enteric_CH4_CO2eq",
                   new_names = "Methane Emissions from Enteric Fermentation")

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation$y$milk_yield,
                        out_var = GHG_simulation$y$enteric_CH4_CO2eq)
```

@@

Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4_CO2eq`.

```{r name_in_out_vars}
in_var  <- GHG_simulation$y$milk_yield
out_var <- GHG_simulation$y$enteric_CH4_CO2eq
```

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4),
                                 max_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4*3))
```
